---
title: "Assignment 1 - Dating"
author: "Shuaitao Tan"
date: "2024-02-14"
output: html_document
---

```{r data import, include=FALSE,message=FALSE,warning=FALSE}
library(readr)
hcmst_data <- read_rds("/Users/shuaitaotan/Documents/GitHub/course_content2/Exercises/03_dating_GRADED/HCMST_couples.rds")
```

```{r, echo=FALSE,message=FALSE,warning=FALSE}
knitr::include_graphics("/Users/shuaitaotan/Tan_Shuaitao_homework/data_v")
```

**Question 1**

```{r Question1 Modify meeting_type, echo=FALSE,message=FALSE,warning=FALSE}
library(ggplot2)
library(dplyr)

hcmst_data <- hcmst_data %>%
  mutate(Simplified_Meeting_Type = case_when(
    meeting_type %in% c("Primary or Secondary School", "College") ~ "School/College",
    meeting_type %in% c("Military", "Customer-Client Relationship", "Work Neighbors", "Business Trip", "Work") ~ "Work-Related",
    meeting_type %in% c("Bar or Restaurant", "Public Place", "Blind Date", "Private Party", "On Vacation", "One-time Service Interaction") ~ "Social/Leisure",
    meeting_type %in% c("Internet", "Internet Dating or Phone App", "Internet Social Network", "Online Gaming", "Internet Chat", "Internet Site", "Met Online") ~ "Online",
    TRUE ~ "Other"
  ))
table(hcmst_data$Simplified_Meeting_Type)
```

```{r Question1 Graph 1,echo=FALSE,message=FALSE,warning=FALSE}
library(ggplot2)
library(dplyr)
# Convert 
hcmst_data <- hcmst_data %>%
  mutate(Q21A_Year = as.numeric(as.character(Q21A_Year))) %>%
  filter(!is.na(Q21A_Year)) %>%
  mutate(Year_Interval = 5 * floor((Q21A_Year - 1930) / 5) + 1930)

# Count 
frequency_data <- hcmst_data %>%
  group_by(Year_Interval, Simplified_Meeting_Type) %>%
  summarise(Frequency = n(), .groups = 'drop')

ggplot(frequency_data, aes(x = Year_Interval, y = Frequency, color = Simplified_Meeting_Type)) +
  geom_line() +
  geom_point() + 
  scale_x_continuous(breaks = seq(1930, 2010, by = 10)) + 
  labs(title = "Trend of Meeting Types Over 5-Year Intervals",
       x = "Year Interval",
       y = "Frequency",
       color = "Meeting Type") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```


I had modified both time interval and the meeting_type. I use trend line to graph out the frequency because it is the easiest way to display change within 5 years of time interval. It shows the change of each category while allowing us to compare the graphs. I recommend this data graph for showing clear trend.

```{r Question1 Graph 2, echo=FALSE,message=FALSE,warning=FALSE}

ggplot(frequency_data, aes(x = factor(Year_Interval), y = Frequency, fill = Simplified_Meeting_Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_x_discrete(breaks = seq(1930, 2010, by = 10), labels = seq(1930, 2010, by = 10)) +
  labs(title = "Frequency of Meeting Types Over 5-Year Intervals",
       x = "Year Interval",
       y = "Frequency",
       fill = "Meeting Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 65, hjust = 1))
```

Here is another visualization of the change, it gives a comparison for frequency between the 5 types of meeting_type. I chose bar chart because it is able to easily compare across categories.



**Question 2**
```{r Question2 Age,echo=FALSE,message=FALSE,warning=FALSE }
selected_variable <- hcmst_data[, c('ppage', 'ppgender', 'Q9')]
# Replace NA in ppage with its mean
selected_variable$ppage[is.na(selected_variable$ppage)] <- mean(selected_variable$ppage, na.rm = TRUE)
# Replace NA in Q9 with its mean
selected_variable$Q9[is.na(selected_variable$Q9)] <- mean(selected_variable$Q9, na.rm = TRUE)
head(selected_variable, 10)


library(ggplot2)
library(ggalt)
updated_colors <- c("Male" = "#9B59B6",  
                    "Female" = "#F1C40F") 

# Basic scatter plot with points colored by gender
p <- ggplot(selected_variable, aes(x = ppage, y = Q9)) +
  geom_point(aes(color = ppgender), alpha = 0.2) +  
  scale_color_manual(values = updated_colors) +
  labs(title = "Relationship Between Respondent's and Partner's Ages by Gender",
       x = "Respondent's Age",
       y = "Partner's Age",
       color = "Gender") +
  theme_minimal() +
  geom_smooth(method = "lm", se = FALSE, color = "black") +  # Add trend line
  theme(legend.position = "bottom")
# Add annotation
p <- p + annotate("text", x =100 , y = 20, label = "Positive relationship between respondent's age and partner's age regardless of gender", size = 2, hjust = 1, color = "black")
print(p)
```


Here we can see a clear positive relationship for age similarity.




**Question 3**
```{r Question3,echo=FALSE,message=FALSE,warning=FALSE}
# Define the mapping for both variables
level_to_numeric <- c(
  "Strong Republican" = -3, 
  "Not Strong Republican" = -2, 
  "Leans Republican" = -1,
  "Undecided/Independent/Other" = 0, 
  "Leans Democrat" = 1, 
  "Not Strong Democrat" = 2, 
  "Strong Democrat" = 3
)
# Ensure 'partyid7' and 'w6_q12' are factors
hcmst_data$partyid7 <- factor(hcmst_data$partyid7)
hcmst_data$w6_q12 <- factor(hcmst_data$w6_q12)

# Apply the mapping to convert levels to numeric values, excluding levels not in the mapping
hcmst_data$partyid7_numeric <- ifelse(hcmst_data$partyid7 %in% names(level_to_numeric), 
                                      as.numeric(level_to_numeric[as.character(hcmst_data$partyid7)]), 
                                      NA)  # Assign NA to levels not in the #mapping

hcmst_data$w6_q12_numeric <- ifelse(hcmst_data$w6_q12 %in% names(level_to_numeric), 
                                    as.numeric(level_to_numeric[as.character(hcmst_data$w6_q12)]), 
                                    NA)  
hcmst_data$difference_numeric <- ifelse(!is.na(hcmst_data$partyid7_numeric) & !is.na(hcmst_data$w6_q12_numeric),
                                        hcmst_data$partyid7_numeric - hcmst_data$w6_q12_numeric,
                                        NA)  # Assign NA where either value is NA
library(ggplot2)
ggplot(hcmst_data, aes(x = difference_numeric, y = duration)) +
  geom_point(alpha = 0.7, color = "#9b59b6", size = 3) +
  theme_minimal(base_size = 20) + 
  labs(title = "Difference vs. Duration",
       x = "Difference in Political Affiliation",  
       y = "Duration") +
  theme(plot.title = element_text(face = "bold", size = 30),
        plot.subtitle = element_text(size = 14),  
        axis.title.x = element_text(size = 16),  
        axis.title.y = element_text(size = 16), 
        axis.text.x = element_text(size = 12),  
        axis.text.y = element_text(size = 12)) 
```



From the graph, we can see, the smaller the difference, the longer the duration is. Therefore, it is clear the more similar couples are in their political affiliation, the stronger their relationship is.Here is a revised attempt: I calculated the absolute difference between political affiliation and then used a heat map, and it is clear that the smaller the difference, the longer duration is.

**Question 4**

```{r question4graph1,echo=FALSE,message=FALSE,warning=FALSE}
# Assign easy numeric values to each level
level_to_numeric_Q17C <- c(
  "Refused" = NA,
  "I am sexually attracted only to men" = 1,
  "I am mostly sexually attracted to men, less often sexually attracted to women" = 2,
  "I am equally sexually attracted to men and women" = 3,
  "I am mostly sexually attracted to women, less often sexually attracted to men" = 4,
  "I am sexually attracted only to women" = 5
)

# Replace the levels with numeric values
hcmst_data$Q17C_numeric <- level_to_numeric_Q17C[hcmst_data$Q17C]

library(ggplot2)
# Define a custom color palette with less bright colors
my_colors <- c("1" = "#5E4FA2", "2" = "#66C2A5", "3" = "#FC8D62", "4" = "#8DA0CB", "5" = "#E78AC3")

# Create a scatter plot
ggplot(hcmst_data, aes(x = Q17C_numeric, y = duration, color = as.factor(Q17C_numeric))) +
  geom_jitter(width = 0.3, alpha = 0.6) +  # Add jitter to avoid overplotting
  scale_color_manual(values = my_colors) +  # Use custom color palette
  labs(title = "Relationship Duration by Sexual Orientation",
       x = "Sexual Orientation",
       y = "Duration") +
  scale_x_continuous(breaks = 1:5, labels = c("1" = "only to men",
                                              "2" = "mostly to men, less to women",
                                              "3" = "equally",
                                              "4" = "mostly to women, less to men",
                                              "5" = "only to women")) +  # Set x-axis labels
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",  
        legend.box = "horizontal", 
        legend.key.height = unit(0.5, "cm"),  
        legend.text = element_text(size = 10)) +  
  guides(color = guide_legend(nrow = 2))
```

Here, we can see, sexual orientation does have influence on the duration of relationship: it seems traditional relationship have longer duration on average. However it may be that LGBTQ are minority which means a lack of data collection. This visualization is good to display density of the data by groups or other categorical data.

```{r question 4 graph 2, echo=FALSE,message=FALSE,warning=FALSE}
breaks <- c(-Inf, 3, 9, 13, Inf)  
labels <- c("Less than High School", "High School Graduate", "Some College", "Bachelor's Degree or Higher")  
# Create a new variable "parental_education" by grouping levels of "Q14"
hcmst_data$parental_education <- cut(as.integer(hcmst_data$Q14), breaks = breaks, labels = labels, include.lowest = TRUE)
levels(hcmst_data$parental_education)

level_to_numeric <- c(
  "Less than High School" = 1,
  "High School Graduate" = 2,
  "Some College" = 3,
  "Bachelor's Degree or Higher" = 4
)
# Convert
hcmst_data$parental_education_numeric <- as.numeric(level_to_numeric[as.character(hcmst_data$parental_education)])
library(ggplot2)

ggplot(hcmst_data, aes(x = parental_education_numeric, fill = duration)) +
  geom_histogram(bins = 10, color = "#5E4FA2", fill = "#f1c40f", alpha = 0.5) +  
  geom_density(aes(y = ..count.., group = 1), color = "black", size = 1) + 
  labs(title = "Distribution of Relationship Duration by Parental Education",
       x = "Parental Education Numeric",
       y = "Duration") +
  scale_fill_brewer(palette = "Set1") + 
  theme_minimal() +
  theme(legend.position = "right") 

```

Using a histogram, we can clearly find that parental education positively correlated to the duration of relationship. The data are skewed to the right which indicates that parental education with some college degree have tends to have longer duration.

**Question 5**

```{r Question 5 graph 1,echo=FALSE,message=FALSE,warning=FALSE}
library(ggplot2)
library(plotly)


gg <- ggplot(hcmst_data, aes(x = parental_education_numeric, fill = duration)) +
  geom_histogram(bins = 10, color = "#5E4FA2", fill = "#f1c40f", alpha = 0.5) +  
  geom_density(aes(y = ..count.., group = 1), color = "black", size = 1) + 
  labs(title = "Distribution of Relationship Duration by Parental Education",
       x = "Parental Education Numeric",
       y = "Duration") +
  scale_fill_brewer(palette = "Set1") + 
  theme_minimal() +
  theme(legend.position = "right")

# Convert ggplot to interactive 
gg_plotly <- ggplotly(gg)
gg_plotly
```

I made the graph 2 from Question 4 interactive using ggplotly.

```{r question 5 graph 2,echo=FALSE,message=FALSE,warning=FALSE}
library(highcharter)
# Ensure 'duration' is numeric
hcmst_data$duration <- as.numeric(hcmst_data$duration)

# Create an interactive scatter plot with highcharter
hc <- hchart(hcmst_data, "scatter", hcaes(x = difference_numeric, y = duration)) %>%
  hc_plotOptions(scatter = list(marker = list(radius = 5, symbol = "circle", states = list(hover = list(enabled = TRUE, lineWidthPlus = 0))))) %>%
  hc_tooltip(pointFormat = "Duration: {point.y}<br/>Difference: {point.x}") %>%
  hc_xAxis(title = list(text = "Difference in Political Affiliation")) %>%
  hc_yAxis(title = list(text = "Duration")) %>%
  hc_legend(enabled = FALSE)  # Disable legend
hc
```

My second choice of making the graph interactive is using highcharter package.This graph is from Question 3 graph 1.

**Question6**

```{r question 6,echo=FALSE}
library(DT)
# Select useful variables and rename them
subset_data <- hcmst_data[, c("Q14", "Q17B", "Q19")]
colnames(subset_data) <- c("Mother Education", "Marriage Number", "Living Status Together")

# Create interactive data table
datatable(
  subset_data,
  options = list(
    pageLength = 10,
    filter = 'top',
    ordering = TRUE
  )
)
```
